Symbolic Weighted Language Models and Automated Music Transcription

### 2021-05-04 main.tex
intro
abstract

\documentclass[a4paper,11pt]{article}
%\setcounter{page}{1}

@@

%% theorem envs
\usepackage{theorem}
\newtheorem{theorem}{Theorem} %[section]
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
\theorembodyfont{\slshape}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
{\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}

@@

% Parsing over infinite alphabet as optimal alignment computation 
% as edit-distance between string and language
% 
\title{Symbolic Weighted Parsing and\\ Automated Music Transcription}
%\title{Weighted Visibly Pushdown Automata\\ Automated Music Transcription}
\author{Florent Jacquemard}
%\institute{INRIA \& CNAM, Paris, France\\
%\email{florent.jacquemard@inria.fr}}

%\titlerunning{WVPA \& AMT}
%\authorrunning{Florent Jacquemard}

@@

%\begin{abstract}
%Symbolic Weighted (SW) extension of symbolic automata...
%\end{abstract}

%the problem: quantitative parsing or symbolic parsing
%of parsing of words over infinite input alphabet.

Weighted parsing~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing} 
consists in analyzing a word in input \wrt a weighted language model, \eg a probabilistic context-free grammar (CFG).
In such a model, a weight value is associated to each transition rule, % production rule
and the rule's weights can be combined with a associative product operator $\otimes$ into 
the weight of a derivation (or \emph{abstract syntax tree}, \ie the result of an analyze).
A second operator $\oplus$, associative and commutative, 
is used to handle ambiguity of the model, 
by summing the weights of the possibly several (in general exponentially many) 
syntax trees associated to a given input word.
Intuitively, $\oplus$ aims at selecting, or ranking, the syntax trees.
The weight domain, equipped with these two operators is assumed, at minima, 
to form a \emph{semiring} $\Semiring$, %\cite{Eilenberg74automata}
such as a min-plus algebra of domain $\mathbb{R} \cup \{ +\infty\}$, 
where $\oplus$ is min and $\otimes$ is plus.
%by ranking 
%making the weight domain a semiring.
Some efficient specialized parsing algorithms~\cite{Huang05kbest} have been proposed in this context 
in order to compute the $n$ best syntax trees of a given input word without having to enumerate them all.
Generally based on dynamic programming, these algorithms rely on 
additional algebraic properties of~$\Semiring$.
-- see \eg~\cite{Huang05kbest} for some NLP applications.
%The extraction of $n$ best list is useful

@@

We present here some weighted language models
%model of weighted CFG 
computing on words over infinite input alphabets %sets of terminal symbol
- but with finite sets of states and transitions rules. 
% non-terminals and production rules).
In their transition rules, input symbols appear as variables %(parameters)
and the weight associated to a transition rule is a function of these variables.
% 
This approach is close to the case of Symbolic Automata (SA)~\cite{dAntoni21CACM}, 
except that the domain for weight values is not restricted to be Boolean, 
like for the guards in the rules of SA, 
but can be an arbitrary commutative semiring (assuming some restrictions).
%
The models defined here are finite automata (\SWA), 
transducers (\SWT) and pushdown automata
with a visibly restriction~\cite{AlurMadhusudan09nested} (\SWVPA).
%wrt an edit distance
The latter model of automata computes on \emph{nested words}~\cite{AlurMadhusudan09nested}, 
a structured form of words parenthesized with markup symbols, 
corresponding to a linearization of trees.
In expressiveness, they correspond to a strict restriction
of weighted CFG. 
%and can be used in a general approach for parsing over infinite input alphabets.
Let $A$ be a \SWVPA, associating $A(w) \in \Semiring$ 
to a given a nested word $w$ (representing a parse tree),
and let a \SWT compute a distance $d$, in $\Semiring$, 
between 2 strings over respectively an infinite input alphabet and the 
same (infinite) alphabet of $A$.
Then, the problem of Symbolic Weighted Parsing is, 
given an input string $s$, to find a nested word $w$ minimizing 
(according to the ranking defined by $\oplus$)
the distance $d(s, w) \otimes A(w)$ between $s$ and $A$, 
as defined in~\cite{Mohri03EDWA}.
%following

@@

In Section~\ref{section:SWA} we introduce \SWA and \SWT, 
then \SWVPA are defined in Section~\ref{section:SWPDA}, 
where a polynomial 1-best algorithm is described that can be use to solve
Symbolic Weighted Parsing.
Finally, \label{sec:transcription}, we present an application 
of this approach to automated music transcription that has been implemented.

@@

We follow the approach of~\cite{Mohri03EDWA} for the computation of distances
between words with weighted transducers, and propose models of weighted automata 
and weighted transducers handling infinite alphabets.

These models generalize weighted automata and transducers over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
by labeling each transition with a weight functions that takes the 
input and output symbols as parameters, instead of a simple weight value.
These functions are similar to the guards of symbolic automata~\cite{Veanes13ciaa}, 
but they can return values in an arbitrary semiring, 
where the latter guards are restricted to the Boolean semiring.

@@

\section{SW Visibly Pushdown Automata}
\label{section:SWVPA}\label{sec:SWVPA}
The following model generalizes Symbolic VPA~\cite{DAntonyAlur14SVPDA}
from Boolean semirings to arbitrary semiring weight domains.

@@

\subsection{Properties}
The class of \SWVPA is closed under the binary operators of the underlying semiring.
%
\begin{proposition}
Let $A_1$ and $A_2$ be two (\SWVPA)
over the same $\Sigma$ and $\Semiring$.
There exists two $\SWVPA$ $A_1 \oplus A_2$ and $A_1 \otimes A_2$, 
effectively constructible, 
such that for all $s \in \Sigma^*$, 
$(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and 
$(A_1 \otimes A_2)(s) = A_1(s) \otimes A_2(s)$.
\end{proposition}
The construction is essentially the same 
as the one of~\cite{DAntonyAlur14SVPDA}, in the case of the Boolean semiring.

>>>>

\documentclass[runningheads]{llncs}
%\documentclass[a4paper,11pt]{article}
%\setcounter{page}{1}

@@

%% theorem envs
\usepackage{theorem}
% \newtheorem{theorem}{Theorem} %[section]
% \newtheorem{definition}[theorem]{Definition}
% \newtheorem{lemma}[theorem]{Lemma}
% \newtheorem{corollary}[theorem]{Corollary}
% \newtheorem{proposition}[theorem]{Proposition}
% \newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
% \theorembodyfont{\slshape}
% \newtheorem{example}[theorem]{Example}
% \newtheorem{remark}[theorem]{Remark}
{\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}

@@

% Parsing over infinite alphabet as optimal alignment computation 
% as edit-distance between string and language
% 
%\title{Symbolic Weighted Parsing and Automated Music Transcription}
\title{Symbolic Weighted Language Models and Automated Music Transcription}
%\title{Weighted Visibly Pushdown Automata\\ Automated Music Transcription}
%\author{Florent Jacquemard}
%\institute{INRIA \& CNAM, Paris, France\\
%\email{florent.jacquemard@inria.fr}}

%\titlerunning{WVPA \& AMT}
\titlerunning{Symbolic Weighted Parsing and Automated Music Transcription}
%\authorrunning{Florent Jacquemard}

@@

\begin{abstract}
%Symbolic Weighted (SW) extension of symbolic automata...

We propose language models called Symbolic Weighted Automata (SWA) 
at the joint %intersection 
between Symbolic Automata (SA) and Weighted Automata (WA), 
as well as Transducers (SWT) and Visibly Pushdown (SWPDA) variants.
%
Like SA, SWA deal with large or infinite input alphabets, 
and like WA, they output a weight value in a semiring domain.
The transitions of WA are labeled by functions from an infinite alphabet into the weight domain. 
This is unlike WA whose transitions are guarded by boolean predicates 
overs symbols in an infinite alphabet 
and also unlike WA whose transitions are labeled by constant weight values, 
and who deal only with finite automata.
%
We present some properties of WA, WT and SWPDA models, 
and show how they can be used to define and solve a variant of parsing 
over infinite alphabets.
%
We also briefly describe the application that motivated the introduction of these models: 
an parse-based approach to automated music transcription.
\end{abstract}

%the problem: quantitative parsing or symbolic parsing
%of parsing of words over infinite input alphabet.
Weighted parsing~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing} 
consists in analyzing a word in input \wrt a weighted language model, \eg a probabilistic context-free grammar (CFG).
In such models, a weight value is associated to each transition rule, % production rule
and the rule's weights can be combined with a associative product operator $\otimes$ into 
the weight of a derivation (or \emph{abstract syntax tree}, \ie the result of an analyze).
A second operator $\oplus$, associative and commutative, 
is moreover used to handle ambiguity of the model, 
by summing the weights of the possibly several (in general exponentially many) 
syntax trees associated to a given input word.
Intuitively, $\oplus$ aims at selecting, or ranking, the syntax trees.
The weight domain, equipped with these two operators is assumed, at minima, 
to form a \emph{semiring} $\Semiring$, %\cite{Eilenberg74automata}
such as a min-plus algebra of domain $\mathbb{R} \cup \{ +\infty\}$, 
where $\oplus$ is min and $\otimes$ is plus.
%by ranking 
%making the weight domain a semiring.
Some efficient specialized parsing algorithms~\cite{Huang05kbest} have been proposed in this context 
in order to compute the $n$ best syntax trees of a given input word without having to enumerate them all.
Generally based on dynamic programming, these algorithms rely on 
additional algebraic properties of~$\Semiring$.
-- see \eg~\cite{Huang05kbest} for some NLP applications.
%The extraction of $n$ best list is useful

@@

We present here some weighted language models
%model of weighted CFG 
computing on words over infinite input alphabets %sets of terminal symbol
- but with finite sets of states and transitions rules. 
% non-terminals and production rules).
In their transition rules, input symbols appear as variables %(parameters)
and the weight associated to a transition rule is a function of these variables.
% 
This approach is close to the case of 
Symbolic Automata (SA)~\cite{dAntoniVeanes17CAV,dAntoni21CACM}, 
except that the domain for weight values is not restricted to be Boolean, 
like for the guards in the rules of SA, 
but can be an arbitrary commutative semiring (assuming some restrictions).
%
The models defined here are finite automata (\SWA),
transducers (\SWT) and pushdown automata
with a visibly restriction~\cite{AlurMadhusudan09nested} (\SWVPA).
%wrt an edit distance
The latter model of automata computes on \emph{nested words}~\cite{AlurMadhusudan09nested}, 
a structured form of words parenthesized with markup symbols, 
corresponding to a linearization of trees.
In expressiveness, they correspond to a strict restriction
of weighted CFG. 
%and can be used in a general approach for parsing over infinite input alphabets.
Let $A$ be a \SWVPA, associating $A(w) \in \Semiring$ 
to a given a nested word $w$ (representing a parse tree),
and let a \SWT compute a distance $d$, in $\Semiring$, 
between 2 strings over respectively an infinite input alphabet and the 
same (infinite) alphabet of $A$.
Then, the problem of Symbolic Weighted Parsing is, 
given an input string $s$, to find a nested word $w$ minimizing 
(according to the ranking defined by $\oplus$)
the distance $d(s, w) \otimes A(w)$ between $s$ and $A$, 
as defined in~\cite{Mohri03EDWA}.
%following

@@

In Section~\ref{section:SWA} we introduce \SWA and \SWT, 
then \SWVPA are defined in Section~\ref{sec:SWVPA}, 
where a polynomial 1-best algorithm is described that can be use to solve
Symbolic Weighted Parsing.
Finally, \label{sec:transcription}, we present an application 
of this approach to automated music transcription that has been implemented.

@@

We follow the approach of~\cite{Mohri03EDWA} for the computation of distances
between words and languages, with weighted transducers, 
extending it to infinite alphabet with models of symbolic weighted automata 
and transducers. % handling infinite alphabets.

These models generalize weighted automata and transducers over finite alphabets, 
see  e.g.~\cite{Mohri03EDWA}, 
by labeling each transition with a weight functions that takes the 
input and output symbols as parameters, instead of a simple weight value.
These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
but they can return values in an arbitrary semiring, 
where the latter guards are restricted to the Boolean semiring.

@@

\section{SW Visibly Pushdown Automata}
\label{section:SWVPA}\label{sec:SWVPA}
The following model generalizes Symbolic VPA~\cite{dAntonyAlur14SVPDA}
from Boolean semirings to arbitrary semiring weight domains.

@@

\subsection{Properties}
The class of \SWVPA is closed under the binary operators of the underlying semiring.
%
\begin{proposition}
Let $A_1$ and $A_2$ be two (\SWVPA)
over the same $\Sigma$ and $\Semiring$.
There exists two $\SWVPA$ $A_1 \oplus A_2$ and $A_1 \otimes A_2$, 
effectively constructible, 
such that for all $s \in \Sigma^*$, 
$(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and 
$(A_1 \otimes A_2)(s) = A_1(s) \otimes A_2(s)$.
\end{proposition}
The construction is essentially the same 
as the one of~\cite{dAntonyAlur14SVPDA}, in the case of the Boolean semiring.
